---
title: "chron_comp_heatmap"
author: "Jori de Leuw"
date: "May 6, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

Load packages
```{r}
library(DESeq2)
library(circlize)
library(RColorBrewer)
library(ComplexHeatmap)
```
Load files with day 0 comparison
```{r}
dir <- "/scratch/bacint/jdeleuw/ananse/influence_data2/chron_comp/"

influence_D0vsD07 <- read.csv2(paste(dir, "day0vsday07.txt", sep = ''), 
                   		header = TRUE, 
                    		sep = '\t',  
                    		stringsAsFactors=FALSE)
influence_D07vsD14 <- read.csv2(paste(dir, "day07vsday14.txt", sep = ''), 
                     		header = TRUE, 
                     		sep = '\t', 
                     		stringsAsFactors=FALSE)
influence_D14vsD21 <- read.csv2(paste(dir, "day14vsday21.txt", sep = ''), 
                      		header = TRUE, 
                      		sep = '\t', 
                      		stringsAsFactors=FALSE)
influence_D21vsD43 <- read.csv2(paste(dir, "day21vsday43.txt", sep = ''), 
                      		header = TRUE, 
                      		sep = '\t', 
                      		stringsAsFactors=FALSE)
influence_D43vsD60 <- read.csv2(paste(dir, "day43vsday60.txt", sep = ''), 
                      		header = TRUE, 
                      		sep = '\t', 
                      		stringsAsFactors=FALSE)
influence_D60vsKC <- read.csv2(paste(dir, "day60vsKC.txt", sep = ''), 
                    		header = TRUE, 
                     		sep = '\t', 
                     		stringsAsFactors=FALSE)
```
Filter TF with minimum influence score:
```{r}
min_Influence_score <- 0.95
comparisons <- list(influence_D0vsD07, influence_D07vsD14, influence_D14vsD21, influence_D21vsD43, influence_D43vsD60, influence_D60vsKC)

res <- lapply(comparisons, function(x) subset(x, sumScaled >= min_Influence_score))
```
Get list of TFs of interest
```{r}
Number <- 0
for (i in 1:length(res)) {
  df <- res[[i]]
  if(Number == 0) {
    TFs_of_interest <- df
    Number <- 1
  }
  else {
    TFs_of_interest <- rbind(TFs_of_interest, df)
  }
}

Filtered_TFs <- sort(unique(TFs_of_interest$factor))
write.table(Filtered_TFs, file = '/scratch/bacint/jdeleuw/ananse/analysis/interesting_TFs.csv', row.names = FALSE, col.names = FALSE)
```
Make data frame with filtered TFs and influence score for each comparison:
```{r}
influence_score <- data.frame(factor = Filtered_TFs,
                              D0vsD7 = 0,
                              D7vsD14 = 0,
                              D14vsD21 = 0,
                              D21vsD43 = 0,
                              D43vsD60 = 0,
                              D60vsKC = 0
			                        )
for (i in 1:length(res)) {
  df <- comparisons[[i]]
  column_index <- i + 1

  influence_score[column_index] <- rev(merge(influence_score, df[c('factor', 'sumScaled')], by = 'factor', all.x = TRUE))[1]
}
#convert NA values to 0
influence_score[is.na(influence_score)] <- 0
#convert data to numeric
influence_score[,2:length(influence_score)] <- as.data.frame(sapply(influence_score[,2:length(influence_score)], as.numeric))
#set row names
row.names(influence_score) <- influence_score$factor
influence_score$factor <- NULL
```
Make heatmap
```{r}
color_influence <- colorRamp2(c(0,0.5,1), c("blue", "white", "red"))

set.seed(100)

hm_influence <- Heatmap(influence_score,
              col = color_influence,
              row_gap= unit(3,"mm"),
              column_gap = unit(3,"mm"),
              column_title = 'TF Influence',
              name = "Influence score",
              row_title = "TFs",
              show_row_names = T,
              clustering_distance_rows = "maximum",                 #based on the maximum absolute deviation
              show_row_dend = FALSE,
              row_dend_reorder = TRUE,
              row_km = 6,
              column_order = order(as.numeric(gsub("column", "", colnames(influence_score))))
              )
# pdf("/scratch/bacint/jdeleuw/ananse/analysis/chron_influence_score_HM.pdf",width=12,height=12,paper='special', useDingbats=FALSE)
draw(hm_influence)
# dev.off
```
Load Count_data:
```{r}
count_data <- read.csv2('/scratch/bacint/jdeleuw/KC_RNA/results/counts/GRCh38-counts.tsv', header = TRUE, sep = '\t', row.names= 1, stringsAsFactors=FALSE)
```
Load column data:
```{r}
col_data <- read.csv('/scratch/bacint/jdeleuw/ananse/expression_data/col_data.tsv', sep = "\t", row.names= 1, header = TRUE)
```

run Deseq2:
```{r}
dds <- DESeqDataSetFromMatrix(countData = count_data,
                              colData = col_data,
                              design = ~ Stage)
dds <- DESeq(dds)
```
Normalize the counts
```{r}
rld <- rlog(dds, blind=FALSE)           #normalize counts
```
Filter counts from TFs of interest
```{r}
list_TFs <- read.csv2("/scratch/bacint/jdeleuw/ananse/analysis/interesting_TFs.csv", 
                            header = FALSE, 
                            sep = '\t',  
                            stringsAsFactors=FALSE)

filtered_counts <- assay(rld)[row.names(assay(rld)) %in% list_TFs$V1 , ]
#z <- mat - rowMeans(mat)
normalized_counts <- data.frame(scale(filtered_counts))
```
Remove two duplicate of KCs and change the column names to more compehensive names:
```{r}
counts <- normalized_counts[, -c(13:14)]
colnames(counts) <- as.matrix(col_data$descriptive_name[-c(13:14)])[,1]
```

```{r}
avg_counts <- data.frame(row.names = row.names(counts),
                      D0 = apply(counts, 1, function(x) mean(as.numeric(x[1:2]))),
                      D7 = apply(counts, 1, function(x) mean(as.numeric(x[3:4]))),
                      D14 = apply(counts, 1, function(x) mean(as.numeric(x[5:6]))),
                      D21 = apply(counts, 1, function(x) mean(as.numeric(x[7:8]))),
                      D43 = apply(counts, 1, function(x) mean(as.numeric(x[9:10]))),
                      D60 = apply(counts, 1, function(x) mean(as.numeric(x[11:12]))),
                      KC = apply(counts, 1, function(x) mean(as.numeric(x[13:14])))
           )
```
Make Heatmap
```{r}
color_counts <- colorRamp2(c(-2,0,2), c("#998ec3","white","#f1a340"))
set.seed(100)
hm_counts <- Heatmap(avg_counts,
              col = color_counts,
              row_gap= unit(3,"mm"),
              column_title = 'Expression',
              name = "Z-score normalized counts",
              row_title = "TFs",
              show_row_names = T,
              clustering_distance_rows = "maximum",                 #based on the maximum absolute deviation
              show_row_dend = FALSE,
              row_dend_reorder = TRUE,
              column_order = order(as.numeric(gsub("column", "", colnames(avg_counts))))
              )
# pdf("/scratch/bacint/jdeleuw/ananse/analysis/chron_TPM_HM.pdf",width=12,height=12,paper='special', useDingbats=FALSE)
draw(hm_counts)
# dev.off
```
merge heatmaps together into one figure:
```{r}
set.seed(100)
hmlist <- hm_influence + hm_counts
pdf("/scratch/bacint/jdeleuw/ananse/analysis/chron_influence_normCounts_HM.pdf",width=12,height=12,paper='special', useDingbats=FALSE)
draw(hmlist,
     column_title = "Comparisons",
     column_title_side = "bottom",
     ht_gap = unit(3, "mm"),
     main = 'Influence score'
     )
dev.off
```
```

