---
title: "Heatmap R script"
author: "Jori de Leuw"
date: "March 11, 2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```
Load packages
```{r}
library(DESeq2)
library(circlize)
library(RColorBrewer)
library(ComplexHeatmap)
```
Load files with day 0 comparison
```{r}
dir <- "/scratch/bacint/jdeleuw/ananse/influence_data/day0_comp/"

influence_D0vsD07 <- read.csv2(paste(dir, "day0vsday07.txt", sep = ''), 
                   		header = TRUE, 
                    		sep = '\t',  
                    		stringsAsFactors=FALSE)
influence_D0vsD14 <- read.csv2(paste(dir, "day0vsday14.txt", sep = ''), 
                     		header = TRUE, 
                     		sep = '\t', 
                     		stringsAsFactors=FALSE)
influence_D0vsD21 <- read.csv2(paste(dir, "day0vsday21.txt", sep = ''), 
                      		header = TRUE, 
                      		sep = '\t', 
                      		stringsAsFactors=FALSE)
influence_D0vsD43 <- read.csv2(paste(dir, "day0vsday43.txt", sep = ''), 
                      		header = TRUE, 
                      		sep = '\t', 
                      		stringsAsFactors=FALSE)
influence_D0vsD60 <- read.csv2(paste(dir, "day0vsday60.txt", sep = ''), 
                      		header = TRUE, 
                      		sep = '\t', 
                      		stringsAsFactors=FALSE)
influence_D0vsKC <- read.csv2(paste(dir, "day0vsKC.txt", sep = ''), 
                    		header = TRUE, 
                     		sep = '\t', 
                     		stringsAsFactors=FALSE)
```
Load files with day 0 comparisons in the other way
```{r}
dir <- "/scratch/bacint/jdeleuw/ananse/influence_data/reverse_day0_comp/"

influence_reverse_D07vsD0 <- read.csv2(paste(dir, "day07vsday0.txt", sep = ''), 
                                    		header = TRUE, 
                                    		sep = '\t',  
                                    		stringsAsFactors=FALSE)
influence_reverse_D14vsD0 <- read.csv2(paste(dir, "day14vsday0.txt", sep = ''), 
                                    		header = TRUE, 
                                    		sep = '\t',  
                                    		stringsAsFactors=FALSE)
influence_reverse_D21vsD0 <- read.csv2(paste(dir, "day21vsday0.txt", sep = ''), 
                                    		header = TRUE, 
                                    		sep = '\t',  
                                    		stringsAsFactors=FALSE)
influence_reverse_D43vsD0 <- read.csv2(paste(dir, "day43vsday0.txt", sep = ''), 
                                    		header = TRUE, 
                                    		sep = '\t',  
                                    		stringsAsFactors=FALSE)
influence_reverse_D60vsD0 <- read.csv2(paste(dir, "day60vsday0.txt", sep = ''), 
                                    		header = TRUE, 
                                    		sep = '\t',  
                                    		stringsAsFactors=FALSE)
influence_reverse_KCvsD0 <- read.csv2(paste(dir, "KCvsday0.txt", sep = ''), 
                                    		header = TRUE, 
                                    		sep = '\t',  
                                    		stringsAsFactors=FALSE)
```

Filter TF with minimum influence score:
```{r}
min_Influence_score <- 0.95
comparisons <- list(influence_D0vsD07, influence_D0vsD14, influence_D0vsD21, influence_D0vsD43, influence_D0vsD60, influence_D0vsKC)
reverse_comparisons <- list(influence_reverse_D07vsD0, influence_reverse_D14vsD0, influence_reverse_D21vsD0, influence_reverse_D43vsD0, influence_reverse_D60vsD0, influence_reverse_KCvsD0)

res <- lapply(comparisons, function(x) subset(x, sumScaled >= min_Influence_score))
res_reverse <- lapply(reverse_comparisons, function(x) subset(x, sumScaled >= min_Influence_score))
```
Get list of TFs of interest
```{r}
total_res <- append(res, res_reverse)

Number <- 0
for (i in 1:length(total_res)) {
  df <- total_res[[i]]
  if(Number == 0) {
    TFs_of_interest <- df
    Number <- 1
  }
  else {
    TFs_of_interest <- rbind(TFs_of_interest, df)
  }
}

Filtered_TFs <- sort(unique(TFs_of_interest$factor))
write.table(Filtered_TFs, file = '/scratch/bacint/jdeleuw/ananse/analysis/interesting_TFs.csv', row.names = FALSE, col.names = FALSE)
```
Make data frame with filtered TFs and influence score for each comparison:
```{r}
influence_score <- data.frame(factor = Filtered_TFs,
                              D0vsD7 = 0,
                              D0vsD14 = 0,
                              D0vsD21 = 0,
                              D0vsD43 = 0,
                              D0vsD60 = 0,
                              D0vsKC = 0
			                        )
rev_influence_score <- data.frame(factor = Filtered_TFs,
                                  D7vsD0 = 0,
                                  D14vsD0 = 0,
                                  D21vsD0 = 0,
                                  D43vsD0 = 0,
                                  D60vsD0 = 0,
                                  KCvsD0 = 0
    			                        )

for (i in 1:length(res)) {
  df <- comparisons[[i]]
  rev_df <- reverse_comparisons[[i]]
  column_index <- i + 1

  influence_score[column_index] <- rev(merge(influence_score, df[c('factor', 'sumScaled')], by = 'factor', all.x = TRUE))[1]
  rev_influence_score[column_index] <- rev(merge(rev_influence_score, rev_df[c('factor', 'sumScaled')], by = 'factor', all.x = TRUE))[1]
}
#convert NA values to 0
influence_score[is.na(influence_score)] <- 0
rev_influence_score[is.na(rev_influence_score)] <- 0
#convert data to numeric
influence_score[,2:length(influence_score)] <- as.data.frame(sapply(influence_score[,2:length(influence_score)], as.numeric))
rev_influence_score[,2:length(rev_influence_score)] <- as.data.frame(sapply(rev_influence_score[,2:length(rev_influence_score)], as.numeric))
#set row names
row.names(influence_score) <- influence_score$factor
influence_score$factor <- NULL

row.names(rev_influence_score) <- rev_influence_score$factor
rev_influence_score$factor <- NULL
```
Make heatmap
```{r}
color_influence <- colorRamp2(c(0,0.5,1), c("blue", "white", "red"))
combined_influence <- cbind(rev_influence_score, influence_score)

set.seed(100)

hm_influence <- Heatmap(combined_influence,
              col = color_influence,
              row_gap= unit(3,"mm"),
              column_gap = unit(3,"mm"),
              column_title = 'TF Influence',
              name = "Influence score",
              row_title = "TFs",
              show_row_names = T,
              clustering_distance_rows = "maximum",                 #based on the maximum absolute deviation
              show_row_dend = FALSE,
              row_dend_reorder = TRUE,
              row_km = 6,
              column_km = 2,
              column_order = order(as.numeric(gsub("column", "", colnames(combined_influence))))
              )
# pdf("/scratch/bacint/jdeleuw/ananse/analysis/influence_score_HM.pdf",width=12,height=12,paper='special', useDingbats=FALSE)
draw(hm_influence)
# dev.off
```
Load Count_data:
```{r}
count_data <- read.csv2('/scratch/bacint/jdeleuw/KC_RNA/results/counts/GRCh38-counts.tsv', header = TRUE, sep = '\t', row.names= 1, stringsAsFactors=FALSE)
```
Load column data:
```{r}
col_data <- read.csv('/scratch/bacint/jdeleuw/ananse/expression_data/col_data.tsv', sep = "\t", row.names= 1, header = TRUE)
```

run Deseq2:
```{r}
dds <- DESeqDataSetFromMatrix(countData = count_data,
                              colData = col_data,
                              design = ~ Stage)
dds <- DESeq(dds)
```
Normalize the counts
```{r}
rld <- rlog(dds, blind=FALSE)           #normalize counts
```
Filter counts from TFs of interest
```{r}
list_TFs <- read.csv2("/scratch/bacint/jdeleuw/ananse/analysis/interesting_TFs.csv", 
                            header = FALSE, 
                            sep = '\t',  
                            stringsAsFactors=FALSE)

filtered_counts <- assay(rld)[row.names(assay(rld)) %in% list_TFs$V1 , ]
#z <- mat - rowMeans(mat)
normalized_counts <- data.frame(scale(filtered_counts))
```
Remove two duplicate of KCs and change the column names to more compehensive names:
```{r}
counts <- normalized_counts[, -c(13:14)]
colnames(counts) <- as.matrix(col_data$descriptive_name[-c(13:14)])[,1]
```

```{r}
avg_counts <- data.frame(row.names = row.names(counts),
                      D0 = apply(counts, 1, function(x) mean(as.numeric(x[1:2]))),
                      D7 = apply(counts, 1, function(x) mean(as.numeric(x[3:4]))),
                      D14 = apply(counts, 1, function(x) mean(as.numeric(x[5:6]))),
                      D21 = apply(counts, 1, function(x) mean(as.numeric(x[7:8]))),
                      D43 = apply(counts, 1, function(x) mean(as.numeric(x[9:10]))),
                      D60 = apply(counts, 1, function(x) mean(as.numeric(x[11:12]))),
                      KC = apply(counts, 1, function(x) mean(as.numeric(x[13:14])))
           )
```
Make Heatmap
```{r}
color_counts <- colorRamp2(c(-2,0,2), c("#998ec3","white","#f1a340"))
set.seed(100)
rowOrder <- unlist(row_order(hm_influence))

set.seed(100)
hm_counts <- Heatmap(avg_counts,
              col = color_counts,
              row_gap= unit(3,"mm"),
              column_title = 'Expression',
              name = "Z-score normalized counts",
              row_title = "TFs",
              clustering_distance_rows = "maximum",                 #based on the maximum absolute deviation
              show_row_dend = FALSE,
              row_order = rowOrder,
              column_order = order(as.numeric(gsub("column", "", colnames(avg_counts))))
              )
# pdf("/scratch/bacint/jdeleuw/ananse/analysis/TPM_HM.pdf",width=12,height=12,paper='special', useDingbats=FALSE)
draw(hm_counts)
# dev.off
```
Make heatmap for the targeting clusters, showing how the TF in each cluster are targeting the other TF in the hm
```{r}
dir <- '/scratch/bacint/jdeleuw/ananse/binding_data/binding_analysis/'
targeting_paths_list <- c(paste(dir, 'targeted_hm_Cluster1.tsv', sep = ''),
                         paste(dir, 'targeted_hm_Cluster2.tsv', sep = ''),
                         paste(dir, 'targeted_hm_Cluster3.tsv', sep = ''),
                         paste(dir, 'targeted_hm_Cluster4.tsv', sep = ''),
                         paste(dir, 'targeted_hm_Cluster5.tsv', sep = ''),
                         paste(dir, 'targeted_hm_Cluster6.tsv', sep = '')
                         )
heatmap_titles <- list('Pluripotency Cluster', 'MYC', 'Initiation Cluster', 'Maturation Cluster', 'Constitutive Cluster', 'Mixed')

for (i in 1:length(targeting_paths_list)) {
  path <- targeting_paths_list[i]
  TF_binding_avg <- read.csv2(path, 
                            header = TRUE,
                            row.names = 1,
                            sep = '\t',  
                            stringsAsFactors=FALSE) 
  #Convert data to numeric
  sorted_binding_avg <- TF_binding_avg[ order(row.names(TF_binding_avg)), ]
  numeric_binding_avg <- as.data.frame(apply(sorted_binding_avg, 2, function(x) as.numeric(x)))
  row.names(numeric_binding_avg) <- row.names(sorted_binding_avg)
  numeric_binding_avg[is.na(numeric_binding_avg)] <- 0
  row_means <- numeric_binding_avg - rowMeans(numeric_binding_avg)
  column_means <- scale(row_means, center = TRUE)
  
  output_name_temp <- strsplit(path, '/')[[1]][length(strsplit(path, '/')[[1]])]
  output_name <- gsub('.tsv', '', output_name_temp)
  
  color_binding <- colorRamp2(c(-0.4,0,0.4), c("purple", "white", "green"))
  # color_binding <- brewer.pal(100, 'Greens')
  set.seed(100)
  rowOrder <- unlist(row_order(hm_influence))

  set.seed(100)
  assign(output_name,
         Heatmap(row_means,
              col = color_binding,
              row_gap= unit(3,"mm"),
              column_title = heatmap_titles[i],
              name = 'average binding score',
              row_title = "TFs",
              show_row_names = T,
              row_order = rowOrder,
              # clustering_distance_rows = "maximum",                 #based on the maximum absolute deviation
              show_row_dend = FALSE,
              # row_dend_reorder = TRUE,
              column_order = order(as.numeric(gsub("column", "", colnames(row_means))))
              )
         )
}
```

merge the HMs:
```{r}
set.seed(100)
hmlist <- hm_influence + hm_counts + targeted_hm_Cluster1 + targeted_hm_Cluster2 + targeted_hm_Cluster3 + targeted_hm_Cluster4 + targeted_hm_Cluster5 + targeted_hm_Cluster6
set.seed(100)
rowOrder <- unlist(row_order(hm_influence))
set.seed(100)
pdf("/scratch/bacint/jdeleuw/ananse/analysis/influence_normCounts_targeted_HM_new_ananse.pdf",width=24,height=12,paper='special', useDingbats=FALSE)
draw(hmlist,
     column_title = "Days",
     column_title_side = "bottom",
     ht_gap = unit(3, "mm"),
     row_order = rowOrder 
     )
dev.off
```
